### 2015-05-12 Dissertation/chapter-conclusions.tex
Updates from Gray-Box GECCO paper

>>>>

In line with previous work, we have found that HBHC cannot effectively find global optima
on problems with even moderate epistasis. In general it also obtains almost no improvement
in fitness after only a few restarts. We designed TUX as a simplistic way of choosing
restart locations based on previously found local optima. Unlike HBHC alone, TUX
was able to continue improving given more time, finding global optima on problems three
times as large.

TUX's effectiveness is likely due to the HBHC acting as a super repair operator for uniform crossover.
Given a sufficiently large $r$ the HBHC can return sections of the crossover offspring to either
parents' original version of a given subfunction. The HBHC is elitist meaning there is a bias toward
returning to the better of the two parent's versions. Furthermore, by being so disruptive,
uniform crossover potentially allows for the HBHC to also find unrelated improvements.

While TUX improves over plain HBHC, Gray-Box P3 is required to perform truly successful
global optimization. Gray-Box P3 replaces na\"{i}ve local search with the HBHC and utilizes
known non-linear relationships instead of statistical linkage learning. On NKq-Landscapes this drastically
improves search effectiveness. A major source of this improvement is likely how difficult it is
for Black-Box P3 to learn linkage relationships on these landscapes. Furthermore, Gray-Box P3
can perform partial reevaluation and efficient hill climbing during the mixing phase.

Gray-Box P3's success is less dramatic on Ising Spin Glasses.
While it still outperformed all competitors, Black-Box P3 may actually scale better to larger
problems. One explanation for this deviation is that Ising Spin Glasses require more
exploration of equal fitness plateaus. For instance in Figure~\ref{fig-solver-over} and Figure~\ref{fig-radius-over}
there is a significant pause in improvement when Gray-Box P3 reaches the second best fitness in the landscape.
Nothing in its design suggests that Gray-Box P3 should be more effective at neutral drift. Another potential
issue is that on these landscapes the importance of each non-linear relationship may be detectably unequal.
As a result Black-Box linkage learning may better cluster variables which having meaningful
impact on fitness while Gray-Box assumes all are equally important. A useful direction for future work
would be to explore methods of performing efficient learning on top of the known variable interactions.

Somewhat surprising is the difference in behavior between the polynomially solvable problems
and Unrestricted NKq. While Black-Box P3 performed very well in the former, it was the least
successful in the latter. The optimal radius for Gray-Box P3 also shifted from 1 to 2.
A potential explanation is that in both Ising Spin Glass
and Nearest Neighbor NKq the number of unique variables $r$
or fewer steps away from a given variable is significantly lower than the worst case. This explains why on Unrestricted NKq
even moderately high $r$ values hit our memory limit.
For Black-Box P3 this may also be causing increased difficulty in linkage learning as variables
become indirectly dependent on much larger sets. Furthermore, Black-Box
P3 may be benefiting from an increased rate of duplicate dependencies on Nearest Neighbor NKq
not present in Unrestricted NKq.

While the inclusion of HBHC into Gray-Box P3 introduces a parameter, it requires
trivial configuration. In the worst case there may be a handful of $r$ values to test.
Furthermore, our evidence suggests setting $r=1$ is quite powerful, with higher values
likely to be only a time shift in quality. This is in contrast to $r$'s role in HBHC,
where low $r$ values are never expected to reach the same quality has higher $r$ values.
Therefore we conclude that Gray-Box P3 maintains the out-of-the-box quality of Black-Box
P3, while drastically improving efficiency for this new domain of problems.
